---
title: "assignment4"
author: "AJ Fowler"
date: "2/17/2020"
output: html_document
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Installations from CRAN
c(  "plotly",         
  "tidyverse",      
  "rvest",
  "ballr",
  "purrr",
  "tseries"        
 
  ) -> package_names  

for(package_name in package_names) {
  if(!is.element(package_name, installed.packages()[,1])) {
     install.packages(package_name,
                      repos = "http://cran.mtu.edu/")
  }
  library(package_name, character.only=TRUE,
          quietly=TRUE,verbose=FALSE)
}

# Installations from devtools. Because of folder structure on GitHub,
#  these must be done individually
if(!is.element("tsdl", installed.packages()[,1])) {
  devtools::install_github("FinYang/tsdl")  # Everything else is on CRAN
}
library(tsdl)

rm(list=c("package_name", "package_names"))

options(show.signif.stars = FALSE)

```
I pulled data from basketball references with the ballR package. I used data from the current NBA season and filtered the data to get the top 50 scorers in the NBA.
```{r data prep}
# reading in data
pergamestats_df <- NBAPerGameStatistics(season = 2020)
advgamestats_df <- NBAPerGameAdvStatistics(season = 2020)

# merging per game and adv stats
merge(pergamestats_df, advgamestats_df, by = "link") -> sznstats_df

# filter for at least 40 games in 2020 season, top 65 pts (to account for duplicates)

sznstats_df%>%
  filter(., gs > 40) %>%
 top_n(n = 65, pts) %>% 
distinct() -> top50scorers_df

# removing duplicates
top50scorers_df[-c(12,13,14,15,16,39,40,41,42,43,60,61,62,63,64),] -> top50scorers_df

# remove extra columns, remaning columns
top50scorers_df%>%
  select(., -rk.x, -g.x, -rk.y, -player.y, -pos.y, -age.y, -tm.y, -g.y, -mp.y, -link) %>%
  rename(player = player.x, pos = pos.x, age = age.x, team = tm.x, mp = mp.x) %>%
  select(., -x, -x_2) -> top50scorers_df
```
 
```{r graphs}
# Top Scorers / Win Shares
scoringplot1 <- plot_ly(top50scorers_df, x = ~pts, y = ~ws,
  # Hover text:
  text = ~paste("Player:",player,"<br>Points Per Game:", pts, "<br>Win Shares:", ws),
  color = ~usgpercent, size = ~bpm) %>%
  # Titles
  layout(title = 'Are the Top Scores in the NBA Effecting Winning',
         yaxis = list(zeroline = TRUE),
         xaxis = list(zeroline = TRUE))
  
scoringplot1

# The top scorers are influencing wining so now my question was what about there scoring is influencing winning

top50scorers_df %>%
  filter(., ws >= 5) -> topwinscorers_df

scoringplot2 <- plot_ly(topwinscorers_df, x = ~ftr, y = ~ws,
  # Hover text:
  text = ~paste("Player:",player,"<br>Win Shares:", ws, "<br>Points Per Game:", pts),
  color = ~usgpercent, size = ~age) %>%
  
  layout(title = 'How are the Best winning players Scoring in the NBA (Free Throws)',
         xaxis = list(title = 'Free Throw Rate'),
         yaxis = list(title = 'Win Shares'))

scoringplot2

# Free Throw Rate seems to be a good indicator of how a scorer influences winning

scoringplot3 <- plot_ly(topwinscorers_df, x = ~x3par, y = ~ws,
  # Hover text:
  text = ~paste("Player:",player,"<br>Win Shares:", ws, "<br>Points Per Game:", pts),
  color = ~usgpercent, size = ~age) %>%
  
  layout(title = 'How are the Best winning players Scoring in the NBA (3 Pointers)',
         xaxis = list(title = '3 Point Rate'),
         yaxis = list(title = 'Win Shares'))

scoringplot3



```
These graphs displayed the top scorers in the NBA, how they score and if their scoring affected winning. In these graphs I quantified winning by the average number of win shares the players have per game. I also included box plus minus in some graphs but Win Shares was the main indicator for wins. I also included usage percentage and you can see a trend in high usage players with higher scoring leading to more win shares. This was probably due to the variables being correlated as usage combines points per game and other variables to get a number for the percentage. So after this I looked into how the top scorers were scoring and if that affected them win shares. Free throw rate didn't really show any trend in terms of if you increase it you will have more of a chance to win but every player with over 4.5% free throw rate was on the higher end of win shares (except Joel Embiid who has one of the lowest win shares of top scorers in the NBA on a good team). Three point rate was more telling in terms of it leading to winning as there is a clear trend. Especially when you remove the big men like Anthony Davis, Giannis, and Jokic who dont take alot of three pointers anyway you can see increasing the three point rate of your superstar scorer could lead to more wins.
